---
title: "Lab 4_Demo"
author: "Mateo Robbins"
date: "2024-04-22"
output: html_document
---

```{r packages, include = FALSE}
library(tidytext)
library(tidyverse)
library(tidymodels)
library(textrecipes)
library(discrim) # naive-bayes
library(naivebayes)
```

This data set includes more possible predictors than the text alone, but for this model we will only use the text variable
```{r data}

urlfile ="https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/climbing_reports_model_dat.csv"
incidents_df<-readr::read_csv(url(urlfile))
```

Now we'll split our data into training and test portions

```{r split-data}
set.seed(1234)

# re-classify deadly column
incidents2class <- incidents_df |> 
  mutate(fatal = factor(if_else(is.na(Deadly),
                                "non-fatal", "fatal")))
# examine split of data; suggests a class imbalance
table(incidents2class$fatal)

# create initial split
incidents_split <- initial_split(incidents2class, strata = fatal)

incidents_train <- training(incidents_split)
incidents_test <- testing(incidents_split)

```

We use recipe() to specify the predictor and outcome variables and the data.

```{r recipe}
incidents_rec <- recipe(fatal ~ Text, data = incidents_train)

recipe <- incidents_rec |> 
  # break text into tokens
  step_tokenize(Text) |> 
  # I only want to use 1000 most frequent words
  step_tokenfilter(Text, max_tokens = 1000) |> 
  # correct for high frequency words that mean little
  # (like stop words and domain specific stop words)
  step_tfidf(Text)

recipe
```

Create  tidymodels workflow to combine the modeling components

```{r workflow}
incidents_wf <- workflow() |> 
  add_recipe(recipe)
```

```{r nb-spec}
nb_spec <- naive_Bayes() |> 
  set_mode("classification") |> 
  set_engine("naivebayes")
```

Now we are ready to add our model to the workflow and fit it to the training data

```{r fit-model}
# initial fit
nb_fit <- incidents_wf |> 
  add_model(nb_spec) |> 
  fit(data = incidents_train)
```

Next up is model evaluation. We'll stretch our training data further and use resampling to evaluate our naive Bayes model. Here we create 10-fold cross-validation sets, and use them to estimate performance.

```{r cv_folds}
set.seed(999)

incidents_folds <- vfold_cv(incidents_train) 

incidents_folds
```

```{r nb-workflow}
nb_wf <- workflow() |> 
  add_recipe(recipe) |> 
  add_model(nb_spec)

nb_wf
```

To estimate its performance, we fit the model to each of the resampled folds, and then evaluate on the heldout part of each resampled fold.

```{r fit-resamples}
nb_rs <- fit_resamples(
  nb_wf,
  incidents_folds,
  control = control_resamples(save_pred = T)
)
```

Extract the relevant information using collect_metrics() and collect_predictions() and examine the performance metrics.

```{r performance}
nb_rs_metrics <- collect_metrics(nb_rs)
nb_rs_predictions <- collect_predictions(nb_rs)

nb_rs_metrics
nb_rs_predictions
```

We'll use two performance metrics: accuracy and ROC AUC.
Accuracy is the proportion of the data that is predicted correctly. 
The ROC curve plots the true positive rate against the false positive rate; AUC closer to 1 indicates a better-performing model, while AUC closer to 0.5 indicates a model that does no better than random guessing.

```{r performance-plot}
nb_rs_predictions |> 
  group_by(id) |> 
  roc_curve(truth = fatal, .pred_fatal) |> 
  autoplot() +
  labs("Resamples", title = "ROC curve for Climbing Incident Reports")
```

Another model method involves the confusion matrix. A confusion matrix tabulates a model's false positives and false negatives for each class.

```{r confusion-matrix}
conf_mat_resampled(nb_rs, tidy = F) |> 
  autoplot(type = "heatmap")
```

*model is good at predicting non-fatal but does so with low-accuracy* 

```{r null-model}

```
